Temperature and moisture effects on greenhouse gas emissions from deep active-layer boreal soils

### 2016-03-30 manuscript/cpcrw_manuscript.Rmd
Added Q10 calculation, corresponding methods & results text

Anomalous data were excluded based on their being more than `r CO2_EXCLUDE_DEVS` (for CO~2~) or `r CH4_EXCLUDE_DEVS` (for CH~4~) mean absolute deviations {Davies, 1993 #3138} from the treatment mean within each day, treatment, and temperature. We excluded `r en` of `r fluxn` (`r round(en/fluxn*100)`%) of the gas flux measurements for this reason. Other data (`r round(rn/sdn*100)`%) were removed because of known instrument problems, e.g. the analyzer was left running after leaving a chamber. The final number of valid flux samples from the 100-day incubation was `r sum(!fluxdata$outlier)`.

The GHG concentration measurements for each sample (i.e., each gas, core, and date/time) were used to calculate a linear rate of change (δc/δt), based on the concentration rise from a minimum (at 2-10 seconds after measurement began) to a maximum (at 10-45 seconds). Each core’s respiration flux (F) was then calculated following e.g. Steduto et al. {, 2002 #3383} as $F=\frac{\delta c}{\delta t} \frac{V}{M} \frac{P_a}{RT}$ where V is the core-specific system volume, M the core dry mass as determined at the end of the incubation, Pa atmospheric pressure (101 kPa; the incubation chambers were ~120 m a.s.l.), R the universal gas constant (8.3 x 10^-3^ m^3^ kPa mol^-1^ K^-1^) and T the chamber air temperature (K) at time of measurement. The final respiration rate was expressed on a dry soil mass basis (mg C kg soil^-1^ day^-1^). TODO check all units

* joined with mass and ancillary data
* correlations
* Q10 computation
* stats

@@

The cumulative flux numbers above imply Q10 values of XXXX...

APS comments to address re above - TODO:

* It doesn’t look like there was an interactive effect of temp*moisture, but just checking if you included that in the stat model?
* This is good IF the controlled drought was really following the same dry-down rate of the 4C drought treatment (see comment on Figure 2 below). This way a comparison between 20C drought and 4C drought is more appropriate and meaningful.
* Totally surprised by this [last sentence above]. Assumed CH4 would be higher in cores sitting in water (i.e. field moist) – perhaps this CH4→ CO2 ‘effect’ is occurring in these cores as well!?!! 
* Cumulative CO2 between cores held at 4˚ C and cores held at 20˚ C in the“field moisture” treatment did not (???!!!) differ (TRUE?!! Kinda looks that way and that is interesting and useful

>>>>

Anomalous data were excluded based on their being more than `r CO2_EXCLUDE_DEVS` (for CO~2~) or `r CH4_EXCLUDE_DEVS` (for CH~4~) mean absolute deviations {Davies, 1993 #3138} from the treatment mean within a 10-day period, treatment, and temperature. We excluded `r en` of `r fluxn` (`r round(en/fluxn*100)`%) of the gas flux measurements for this reason. Other data (`r round(rn/sdn*100)`%) were removed because of known instrument problems, e.g. the analyzer was left running after leaving a chamber. The final number of valid flux samples from the 100-day incubation was `r sum(!fluxdata$outlier)`.

The GHG concentration measurements for each sample (i.e., each gas, core, and date/time) were used to calculate a linear rate of change (δc/δt), based on the concentration rise from a minimum (at 2-10 seconds after measurement began) to a maximum (at 10-45 seconds). Each core’s respiration flux (*F*) was then calculated following e.g. Steduto et al. {, 2002 #3383} as $F=\frac{\delta c}{\delta t} \frac{V}{M} \frac{P_a}{RT}$ where *V* is the core-specific system volume, *M* the core dry mass as determined at the end of the incubation, *P~a~* atmospheric pressure (101 kPa; the incubation chambers were ~120 m a.s.l.), *R* the universal gas constant (8.3 x 10^-3^ m^3^ kPa mol^-1^ K^-1^) and *T* the chamber air temperature (K) at time of measurement. The final respiration rate was expressed on a dry soil mass basis (mg C kg soil^-1^ day^-1^). TODO check all units

Temperature sensitivity (Q~10~) was calculated for each gas and treatment as $\frac{F_2}{F_1}^{\left( \frac{10}{T_2-T_1} \right)}$ where *F~1~* and *F~2~* are the cumulative gas fluxes (mg C) at temperatures *T~1~* and *T~2~* (°C), respectively.

* joined with mass and ancillary data
* correlations

@@

The cumulative flux numbers above result in CH~4~ temperature sensitivity (Q~10~) values of `r pn(q10_cumulative["Field moisture", "CH4"], 1)` and `r pn(q10_cumulative["Drought", "CH4"], 1)` for the field moisture and drought treatments, respectively; the corresponding Q~10~ values  based on cumulative CO~2~ were `r pn(q10_cumulative["Field moisture", "CO2"], 1)` and `r pn(q10_cumulative["Drought", "CO2"], 1)`.
